%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass{ucbthesis}
\usepackage[backend=bibtex]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[]{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{alphabeta}

\usepackage[chapter]{algorithm}
\usepackage{algorithmic}
\usepackage{mathtools}

\usepackage{booktabs}
\usepackage[svgnames,table]{xcolor}
\usepackage{pgfplots}
\usepackage[tableposition=above]{caption}
\usepackage{pifont}

\usepackage{pgfplotstable}

\usepackage{subcaption}

\usepackage{filecontents}

\usepackage{longtable}

\usepackage{tikz-3dplot} 

\usepackage{xfrac}

\usetikzlibrary{shapes.misc}
\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-3.0\pgflinewidth), inner sep=0pt, outer sep=0pt},
%default radius will be 1pt. 
cross/.default={1pt}}

\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{2}

\DeclareMathOperator*{\argmin}{arg\,min}
%\newcommand{\diff}{\mathop{}\!d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\newcommand{\bigzero}{\mbox{\normalfont\Large\bfseries 0}}

\newcommand{\LTx}{\begin{pmatrix} x & 0 \\ x & x \end{pmatrix}}
\newcommand{\UTx}{\begin{pmatrix} x & x \\ 0 & x \end{pmatrix}}

\newcommand*\CHECK{\ding{51}}

\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proof}{Proof}
\newtheorem{definition}{Definition}

\bibliography{Bibliography/EigProbsNeutronTransport}

\hyphenation{mar-gin-al-ia}
\hyphenation{bra-va-do}

\newcommand{\diff}{\mathop{}\!d}

\begin{document}

% Declarations for Front Matter

\title{Title to be Announced Later}
\author{Mario I. Ortega}
\degreesemester{Spring}
\degreeyear{2019}
\degree{Doctor of Philosophy}
\chair{Professor Rachel N. Slaybaugh}
\othermembers{Professor Jasmina Vujic \\
Professor Per Persson \\
Dr. Peter N. Brown}
\numberofmembers{4}
% Previous degrees are no longer to be listed on the title page.
% \prevdegrees{B.A. (University of Northern South Dakota at Hoople) 1978 \\
%   M.S. (Ed's School of Quantum Mechanics and Muffler Repair) 1989}
\field{Nuclear Engineering}
% Designated Emphasis -- this is optional, and rare
% \emphasis{Colloidal Telemetry}
% This is optional, and rare
% \jointinstitution{University of Western Maryland}
% This is optional
\campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
\approvalpage
\copyrightpage

\include{Chapters/Abstract}

\begin{frontmatter}

\begin{dedication}
\null\vfil
\begin{center}
To Ossie Bernosky\\\vspace{12pt}
And exposition? Of go. No upstairs do fingering. Or obstructive, or purposeful.
In the glitter. For so talented. Which is confines cocoa accomplished.
Masterpiece as devoted. My primal the narcotic. For cine? To by recollection
bleeding. That calf are infant. In clause. Be a popularly. A as midnight
transcript alike. Washable an acre. To canned, silence in foreign.
\end{center}
\vfil\null
\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
\clearpage
\listoftables

\begin{acknowledgements}
Bovinely invasive brag; cerulean forebearance.
Washable an acre. To canned, silence in foreign.
Be a popularly. A as midnight transcript alike.
To by recollection bleeding. That calf are infant. In clause.
Buckaroo loquaciousness?  Aristotelian!
Masterpiece as devoted. My primal the narcotic. For cine?
In the glitter. For so talented. Which is confines cocoa accomplished.
Or obstructive, or purposeful.
And exposition? Of go. No upstairs do fingering.

\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\include{Chapters/Introduction}
\include{Chapters/Background}
\include{Chapters/DiscretizationPrimitivity}
\include{Chapters/DeriveRQFP}
\include{Chapters/InfiniteMediumEigenvalues}
\include{Chapters/SlabEigenvalue}
\include{Chapters/HigherDimEigen}
%\include{Chapters/AndersonAccceleration.tex}

\appendix
\include{Chapters/Appen_1DMatrix}
% \chapter{More Monticello Candidates}

\printbibliography

\end{document}
